### 2023-02-08 writing/manuscript.Rmd
updating meta-analysis link, remove three layer model
moving across final updates to public repository

---
header-includes:
- \usepackage[left]{lineno}
- \usepackage{mdframed}
- \usepackage{soul}
- \usepackage{xcolor}
- \definecolor{bleu}{HTML}{2200cc}
- \usepackage{caption}
- \usepackage{ragged2e}
- \usepackage{afterpage}
- \usepackage[labelformat = empty]{caption}
- \usepackage{rotating}
output: 
  pdf_document:
    fig_caption: yes
    number_sections: true
    latex_engine: xelatex
csl: apa.csl
bibliography: "themusiclab.bib"
nocite: "@Bailey1988; @vanderHulst2010; @Inkelas1988; @Chen2016; Ngo2016; @Swaminathan2021; @Choi2021; @Evans2018a; @Jefferies1990; @Niesler2005; @Manyah2006; @Eme2022; @Westermeyer1977; @Hamann2015; @Goldsmith2011; @Ameka2001"
urlcolor: bleu
linkcolor: bleu
---

@@

full_run <- FALSE
if (full_run) {
  source(here("analysis", "subsampling.R"))
  source(here("analysis", "analyses.R"))
  knitr::knit(here("viz", "figures.Rmd"))
}

@@

<!--
Tables and figures...

Fig 1. Sample sizes (word plot)
Fig 2. Main effects summarized w/ music-lesson as reference point
Fig 3. Language-wise forestplots
Table 1. Main effects statistical reporting (mixed models)

SI
Text 1.1 Meta-analysis
Text 1.2 Headphone check
Text 1.3 Preregistered version of everything

Fig S1. Headphone check scores

Table S1. All languages, samples etc
Table S2. Demographics
Table S3. Income-controlled US model
Table S4. Education-controlled US model
Table S5. East v. West model
Table S6A & S6B. Exploratory analysis 
-->

@@

```{=tex}
\begin{mdframed}[backgroundcolor=gray!20]
Tonal languages differ from other languages in their use of pitch (tones) to distinguish words. Lifelong experience speaking and hearing tonal languages has been argued to shape auditory processing in ways that generalize beyond the perception of \textit{linguistic} pitch to the perception of pitch in other domains like \textit{music}. To examine this, we first conducted a meta-analysis whose results were consistent with this idea, but limited by mostly small sample sizes in only a few tonal languages and countries, which made it challenging to disentangle the effects of linguistic experience from variability in music training experience, cultural differences, and other potential confounds. To address these issues, we used web-based citizen science to test this question on a global scale. We assessed music perception skill in $n = `r f(users$speakers$Tonal)`$ native speakers of `r f(users$langs$Tonal)` tonal languages and compared their performance to $n = `r f(users$speakers$other)`$ native speakers of other languages, including `r f(users$langs$Pitch_accented)` pitch-accented and `r f(users$langs$Non_tonal)` non-tonal languages (e.g. Mandarin, Yoruba; Japanese; Hungarian, respectively). Whether or not participants had taken music lessons, native speakers of all 19 tonal languages had an improved ability to discriminate musical melodies. But this improvement came with a trade-off: relative to speakers of pitch-accented or non-tonal languages, tonal language speakers were also worse at processing the musical beat. These results, which held across tonal languages from a variety of geographic regions and were robust to geographic and demographic variation, demonstrate that linguistic experience shapes music perception ability, with implications for relations between music, language, and culture in the human mind.
\end{mdframed}
```

@@

Human perception readily adapts to these soundscapes: early speech experiences tune our hearing to the speech contrasts of our native language(s) [@Kuhl2004; @Werker1984; @Polka1994], and musical experiences during the same time period are thought to have similar "perceptual narrowing" effects, biasing listeners' interpretations of musical rhythm and pitch based on their own musical cultures [@Hannon2005; @Lynch1990]. These effects may cross domains. While music training has minimal causal effects on high-level cognitive skills [@Sala2020; @Mehr2013a], it may sharpen lower-level aspects of speech processing [@Patel2011; @Wong2007] and auditory perception [@Kraus2010]. In the opposite direction, enhanced experience with the types of linguistic pitch used in tonal languages has been argued to shape pitch processing in music [@Bradley2016; @Bidelman2013; @Pfordresher2009].

@@

The lexical use of pitch in tonal languages is distinct from how pitch is otherwise used in speech. For example, many languages use pitch to convey affect [@Cowen2019]; to cue non-lexical meaning [e.g., helping to differentiate between questions and statements\; @Patel2008; @Tong2005]; to emphasize information [@Breen2010]; to cue sentence structure with metrical stress patterns [@Wagner2019], supporting comprehension [@Hilton2021b]; and/or as a cue to speech categories, as in infant- or child-directed speech [@Hilton2022a]. While these many uses of pitch are typical of speech in both tonal and non-tonal languages [e.g., many Indo-European, South Asian, or Australian languages\; @Maddieson2013], in non-tonal languages, pitch is never used lexically to denote word meanings. Lastly, *Pitch-accented* languages is an intermediate category with limited or mixed use of lexical pitch [such as Croatian\; @vanderHulst2011]; note, however, that whether pitch-accented languages form a coherent standalone category or whether they are better considered on a spectrum between tonal and non-tonal languages, with some mixed cases, is a matter of debate [e.g., @Gussenhoven2004; @Hyman2006; @Hyman2009].^[Indeed, a category of languages that groups evidently disparate languages together, such as Japanese and Swedish, may not be defensible. As pitch-accented languages are not our primary focus, here we treat them as a separate group from tonal and non-tonal languages, but also conduct some analyses at the language level rather than the language-type level. <!--We encourage anyone interested in alternative groupings of languages to use our public data and code to re-analyze accordingly.-->]

@@

First, and most importantly, prior studies sample tonal languages narrowly, typically comparing Mandarin or Cantonese speakers from mainland China to English speakers from the United States. Of the tonal language speakers in prior studies, approximately 92% spoke Mandarin or Cantonese; and of the non-tonal speakers, approximately 85% spoke English. As such, no claim about the effects of language experience on music perception on the basis of the prior literature are justifiable, because it is not clear whether prior effects generalize beyond the select few highly studied languages.

Second, the majority of prior studies have very low statistical power due their small sample sizes, which may produce unreliable group-level estimates of effects and increase the risk of bias [@Kraemer1998]. We estimated the power of each prior study to detect effects of $d = 0.5$, a "medium" size effect comparable to the meta-analytic effect estimated for melodic discrimination tasks (Table 1). Across the three categories of music processing tasks, power was low (for studies of melodic discrimination, median power = `r round(median(melodic.gen.random$Power),2)`; for fine-grained pitch discrimination, median power = `r round(median(pitch.gen.random$Power),2)`; for rhythm tasks, median power = `r round(median(rhythm.gen.random$Power),2)`).

@@

Thus, the meta-analysis demonstrates evidence for a potential effect of tonal language experience on melodic discrimination ability and finer-grained pitch discrimination ability — but a lack of linguistic diversity, low statistical power, and high likelihood of systematic bias in the samples studied warrant caution. These issues can be addressed by studying many native speakers of many languages, with and without music training experience, all of whom complete the same assessments of music processing ability. In the rest of the paper, we report a study of the ability to discriminate melodies, detect mistuned singing, and detect misaligned beats, in `r f(users$final)` people across the globe. Participants self-reported their native language, location, demographics, and degree of musical training, enabling language-wise and language-type-wise analyses of each of these musical abilities and their relations to linguistic and musical experience.

@@

Participants were visitors to the citizen-science website <https://themusiclab.org> who completed a set of three music perception tasks presented as an online game (*Test Your Musical IQ*). We did not recruit participants directly; rather, they visited the site after hearing about it organically (e.g., via Reddit posts, YouTube clips, Twitch streams). Participants gave informed consent under an ethics protocol approved either by the Harvard University Committee on the Use of Human Subjects (protocol IRB2017-1206) or the Yale University Human Research Protection Program (protocol 2000033433). Our analysis was preregistered and deviations are noted below. Please see “Data, code, and materials availability” section for more details.

@@

The resulting sample ($N$ = `r f(users$final)`; see Figure 1 and Table S1) included `r f(users$speakers$Tonal)` native speakers of `r f(users$langs$Tonal)` tonal languages; `r f(users$speakers$Pitch_accented)` native speakers of `r f(users$langs$Pitch_accented)` pitch-accented languages; and `r f(users$speakers$Non_tonal)` native speakers of `r f(users$langs$Non_tonal)` non-tonal languages. Languages were primarily classified based on the World Atlas of Language Structures [@Maddieson2013] and the Lyon-Albuquerque Phonological Systems Database [@Maddieson2014]. Languages that are not present in either database were classified according to information from the Phonetics Information Base and Lexicon Database [@Moran2019] or other sources from the linguistics literature. <!--No discrepancies were found across these sources. JL: WALS did classify 3 of the African languages as simple tone, which is unclear whether completely overlap with PA. I added a footnote to this in Table S1. Commenting out this statement unless others think it's warranted--> A summary of all languages studied here, with further classification details and language-wise sample sizes, is in Table S1.

@@

As in the original tasks cited above, each subtask was presented adaptively via `psychTestR` [@Harrison2020]. To minimize the duration of the experiment, we fixed the length of each subtask at 15 trials, the minimum number of trials with acceptably low mean standard errors, according to the original task designs. Demographic items were presented via `jsPsych` [@deLeeuw2015]. The citizen-science platform distributed the experiments using a modified pre-release version of `pushkin` [@Hartshorne2019]. Readers can try this implementation of the three music perception tasks at <https://themusiclab.org/quizzes/miq>.

@@

The random-effect structure is particularly helpful for correcting for sampling imbalances, ensuring that no particular languages or countries can dominate the overall effect, while also allowing us to model variation across languages and countries directly (see, e.g., Figure 3). Note that this analysis approach deviates from our preregistered analysis plan, which involved applying a linear regression model at several sampling levels, without random effects. We made this change on the suggestion of a reviewer, and given the utility of mixed-effects models in measuring cultural variation [e.g., @Hilton2022]. For transparency, we report analyses and results from the preregistered approach in SI Text 1.2; these largely replicate the results reported in the next section.

@@

The results robustly replicated the main findings. Tonal and non-tonal speakers could be reliably distinguished on average for both melodic discrimination ($p$ `r permutation_tests$T_NT$mdt$p %>% fp`) and beat alignment ($p$ `r permutation_tests$T_NT$cabat$p %>% fp`) tasks, but not for the mistuning perception task ($p$ `r permutation_tests$T_NT$mpt$p %>% fp`), thus convergently supporting the reliability of our main results.

Last, we used the same linear discriminant analysis approach to compare pitch-accented to non-tonal language speakers (Supplementary Figure 3). Here, the approach failed to replicate the small advantage to pitch-accent language speakers on the beat alignment task ($p$ `r permutation_tests$PA_NT$cabat$p %>% fp`) identified by our mixed-effect modeling analysis (Table 2), but did find a small advantage for vocal mistuning ($p$ `r permutation_tests$PA_NT$mpt$p %>% fp`). We take this mixed result as cause to not interpret either effect for pitch-accent speakers, given both the less clear theoretical basis for such an effect and the noted ambiguity of the pitch-accented language type [@Hyman2009].

@@

First, we tested whether the socioeconomic status of participants was associated with their task scores, as socioeconomic status may vary systematically with both country and native language in our data, and may mediate the opportunities a person has for musical development. For participants who reported living in the United States, we collected additional demographic information. To assess how socioeconomic status may mediate the main findings, we analysed this subset of participants who reported income information ($n$ = `r f(income_n)`). Despite being restricted to the United States, this sub-sample was nonetheless linguistically diverse, including `r f(users$income$Non_tonal$n)` non-tonal speakers of `r f(users$income$Non_tonal$n_langs)` languages, `r f(users$income$Tonal$n)` tonal speakers of `r f(users$income$Tonal$n_langs)` languages, and `r f(users$income$Pitch_accent$n)` speakers of pitch-accent languages of `r f(users$income$Pitch_accent$n_langs)` languages. We analysed these data using a mixed-effect model of the same structure used in the main analysis, except with an additional fixed-effect term for income. The results show that while income does positively predict performance, such that those with higher incomes tended to perform better, the tonal language melodic discrimination advantage and beat alignment disadvantage held robustly after accounting for these income effects (full statistical reporting is in Table S3).

@@

Our results also help clarify the previously mixed pattern of results concerning the effects of linguistic experience on music processing across different tasks and samples. For example, an advantage for tonal language speakers in melodic pattern processing is consistent with the majority of previous studies [@Alexander2008; @Bradley2016; @Choi2021; @Chen2016; @Swaminathan2021; @Wong2012], though not all [@Peretz2011; @Stevens2013; @Zheng2018]. Meanwhile, similar levels of performance on the fine-grained pitch task between tonal and non-tonal language speakers is supported by some prior studies [@Bent2006; @Bidelman2011a; @Stevens2013; @Tong2018; @Zheng2018] but not other studies that point to a tonal advantage/disadvantage on related tasks [@Bidelman2013; @Chang2016; @Giuliano2011; @Hutka2015; @Peretz2011; @Pfordresher2009]. Lastly, while rhythmic abilities in tonal language speakers have been less studied [see @Wong2012; @Zhang2020], a disadvantage in beat discrimination is consistent with recent work showing that tonal speakers give more weight to pitch cues than duration cues; this weighting cuts across auditory domains [@Jasmin2021]. By leveraging a consistent set of tasks and a large sample size, our results make clear that speaking a tonal language has a measurable connection to music perception ability — but not a uniform one, as this linguistic experience produces both positive and negative effects on perceptual abilities of different types.

Why might tonal language experience have these specific effects on music perception? The answer may lie in the shared mechanisms and neural processing resources associated with auditory perception, whether applied to language or music [@Patel2011; @Patel2008; @Kraus2010; @Asaridou2013; @Peretz2015; although see also @Asano2021; @Bidelman2013]. Both tonal languages and music rely on specialized pitch-based sound categories (tone contours or levels in speech; pitch motifs and melodies in music). If these categories are learned and processed through shared, domain-general mechanisms, then improving the efficiency of these mechanisms through practice in either domain should result in mutual improvements [@Asaridou2013 ;@Chang2016; @Delogu2006; @Delogu2010; @Patel2008; @McMullen2004]. This idea highlights differences between putatively domain-general auditory processing and putatively domain-specific mechanisms that underlie music processing [e.g., the processing of pitch in the context of a tonal hierarchy\; @Albouy2020; @Krumhansl2004; @Peretz2003; @Zatorre2002].

Our results do not explain, however, *how* these mechanisms might be improved by experience. One possibility is that language experience could shape domain-general perceptual strategies regarding inferences about high-level perceptual categories on the basis of low-level cues: acquired perceptual biases (i.e., from tonal language experience) may aid the processing of some stimuli while worsening the processing of others. In speech, listeners give more perceptual weight to cues that are more informative in discriminating contrasts that are salient in their native language [@Schertz2020; @Jasmin2021] and tonal language speakers rely more heavily on pitch to categorize and produce speech stress when acquiring a non-tonal L2 language compared to native speakers [@Wang2008; @Yu2010].

@@

From the mean and standard deviation, we calculated Hedge's *G* for each task and used it as our effect size measure. We then ran a random-effects model for each category of the tasks. \hl{Please refer to XX for commented code of the analysis.}

@@

```{r FigureS3}
#| fig.cap = "\\textbf{Figure S3.} To test whether speakers of pitch-accent and non-tonal languages could be distinguished on the basis of their music perception scores, we ran a permuted Discriminant Function Analysis (pDFA). The three histograms show the null distributions for each music perception task, representing the ability to discriminate 'tonal' from 'non-tonal' when these labels have been randomly shuffled, resulting from 10,000 permutations. The dotted vertical line then represents the actual discrimination performance on non-shuffled data. And the p-value is computed as a approximate test from this null distribution and actual score. "
knitr::include_graphics(here("viz", "figures", "FigureS3-1.pdf"))
```

>>>>

---
header-includes:
- \usepackage[left]{lineno}
- \usepackage{mdframed}
- \usepackage{soul}
- \usepackage{xcolor}
- \definecolor{bleu}{HTML}{2200cc}
- \usepackage{caption}
- \usepackage{ragged2e}
- \usepackage{afterpage}
- \usepackage[labelformat = empty]{caption}
- \usepackage{rotating}
output: 
  pdf_document:
    fig_caption: yes
    number_sections: true
    latex_engine: xelatex
csl: apa.csl
bibliography: "citations.bib"
nocite: "@Bailey1988; @vanderHulst2010; @Inkelas1988; @Chen2016; Ngo2016; @Swaminathan2021; @Choi2021; @Evans2018a; @Jefferies1990; @Niesler2005; @Manyah2006; @Eme2022; @Westermeyer1977; @Hamann2015; @Goldsmith2011; @Ameka2001"
urlcolor: bleu
linkcolor: bleu
---

@@

full_run <- FALSE
if (full_run) {
  source(here("analysis", "analyses.R"))
  knitr::knit(here("viz", "figures.Rmd"))
}

@@

@@

```{=tex}
\begin{mdframed}[backgroundcolor=gray!20]
Tonal languages differ from other languages in their use of pitch (tones) to distinguish words. Lifelong experience speaking and hearing tonal languages has been argued to shape auditory processing in ways that generalize beyond the perception of \textit{linguistic} pitch to the perception of pitch in other domains like \textit{music}. To examine this, we first conducted a meta-analysis, finding moderate evidence for this idea, but in studies strongly limited by mostly small sample sizes in only a few tonal languages and countries. This makes it challenging to disentangle the effects of linguistic experience from variability in music training experience, cultural differences, and other potential confounds. To address these issues, we used web-based citizen science to test this question on a global scale. We assessed music perception skill in $n = `r f(users$speakers$Tonal)`$ native speakers of `r f(users$langs$Tonal)` tonal languages (e.g., Mandarin, Yoruba) and compared their performance to $n = `r f(users$speakers$other)`$ native speakers of other languages, including `r f(users$langs$Pitch_accented)` pitch-accented (e.g., Japanese) and `r f(users$langs$Non_tonal)` non-tonal languages (e.g., Hungarian). Whether or not participants had taken music lessons, native speakers of all 19 tonal languages had an improved ability to discriminate musical melodies. But this improvement came with a trade-off: relative to speakers of pitch-accented or non-tonal languages, tonal language speakers were also worse at processing the musical beat. These results, which held across tonal languages from a variety of geographic regions and were robust to geographic and demographic variation, demonstrate that linguistic experience shapes music perception ability, with implications for relations between music, language, and culture in the human mind.
\end{mdframed}
```

@@

Human perception readily adapts to these soundscapes: early speech experiences tune our hearing to the speech contrasts of our native language(s) [@Kuhl2004; @Werker1984; @Polka1994], and musical experiences during the same time period are thought to have similar "perceptual narrowing" effects, biasing listeners' interpretations of musical rhythm and pitch based on their own musical cultures [@Hannon2005; @Lynch1990]. These effects may cross domains. While music training has minimal causal effects on high-level cognitive skills [@Sala2020; @Mehr2013], it may sharpen lower-level aspects of speech processing [@Patel2011; @Wong2007] and auditory perception [@Kraus2010]. In the opposite direction, enhanced experience with the types of linguistic pitch used in tonal languages has been argued to shape pitch processing in music [@Bradley2016; @Bidelman2013; @Pfordresher2009].

@@

The lexical use of pitch in tonal languages is distinct from how pitch is otherwise used in speech. For example, many languages use pitch to convey affect [@Cowen2019]; to cue non-lexical meaning [e.g., helping to differentiate between questions and statements\; @Patel2008; @Tong2005]; to emphasize information [@Breen2010]; to cue sentence structure with metrical stress patterns [@Wagner2019], supporting comprehension [@Hilton2021b]; and/or as a cue to speech categories, as in infant- or child-directed speech [@Hilton2022a]. While these many uses of pitch are typical of speech in both tonal and non-tonal languages [e.g., many Indo-European, South Asian, or Australian languages\; @Maddieson2013], in non-tonal languages, pitch is never used lexically to denote word meanings. Lastly, *pitch-accented* languages form an intermediate category with limited or mixed use of lexical pitch [such as Croatian\; @vanderHulst2011]; note, however, that whether pitch-accented languages form a coherent standalone category or whether they are better considered on a spectrum between tonal and non-tonal languages, with some mixed cases, is a matter of debate [e.g., @Gussenhoven2004; @Hyman2006; @Hyman2009].^[Indeed, a category of languages that groups evidently disparate languages together, such as Japanese and Swedish, may not be defensible. As pitch-accented languages are not our primary focus, here we treat them as a separate group from tonal and non-tonal languages, but also conduct some analyses at the language level rather than the language-type level.]

@@

First, and most importantly, prior studies sample tonal languages narrowly, typically comparing Mandarin or Cantonese speakers from mainland China to English speakers from the United States. Of the tonal language speakers in prior studies, approximately 92% spoke Mandarin or Cantonese; and of the non-tonal speakers, approximately 85% spoke English. As such, no claim about the effects of language experience on music perception on the basis of the prior literature is justifiable, because it is not clear whether prior effects generalize beyond a few frequently studied languages.

Second, the majority of prior studies have low statistical power, due their small sample sizes, which may produce unreliable group-level estimates of effects and increase the risk of bias [@Kraemer1998]. We estimated the power of each prior study to detect effects of $d = 0.5$, a "medium" size effect comparable to the meta-analytic effect estimated for melodic discrimination tasks (Table 1). Across the three categories of music processing tasks, power was low (for studies of melodic discrimination, median power = `r round(median(melodic.gen.random$Power),2)`; for fine-grained pitch discrimination, median power = `r round(median(pitch.gen.random$Power),2)`; for rhythm tasks, median power = `r round(median(rhythm.gen.random$Power),2)`).

@@

Thus, the meta-analysis demonstrates evidence for a potential effect of tonal language experience on melodic discrimination ability and finer-grained pitch discrimination ability — but a lack of linguistic diversity, low statistical power, and high likelihood of systematic bias in the samples studied warrant caution. These issues can be addressed by studying many native speakers of many languages, with and without music training experience, all of whom complete the same assessments of music processing ability.

In the rest of this paper, we report a study of the ability to discriminate melodies, detect mistuned singing, and detect misaligned beats, in `r f(users$final)` people across the globe. Participants self-reported their native language, location, demographics, and degree of musical training, enabling language-wise and language-type-wise analyses of each of these musical abilities and their relations to linguistic and musical experience.

@@

Participants were visitors to the citizen-science website <https://themusiclab.org> who completed a set of three music perception tasks presented as an online game (*Test Your Musical IQ*). We did not recruit participants directly; rather, they visited the site after hearing about it organically (e.g., via Reddit posts, YouTube clips, Twitch streams). Participants gave informed consent under an ethics protocol approved either by the Harvard University Committee on the Use of Human Subjects (protocol IRB2017-1206) or the Yale University Human Research Protection Program (protocol 2000033433). Our analysis was preregistered and deviations are noted below. See "Data, code, and materials availability" for more details.

@@

The resulting sample ($n$ = `r f(users$final)`; see Figure 1 and Table S1) included `r f(users$speakers$Tonal)` native speakers of `r f(users$langs$Tonal)` tonal languages; `r f(users$speakers$Pitch_accented)` native speakers of `r f(users$langs$Pitch_accented)` pitch-accented languages; and `r f(users$speakers$Non_tonal)` native speakers of `r f(users$langs$Non_tonal)` non-tonal languages.

Languages were primarily classified based on the World Atlas of Language Structures [@Maddieson2013] and the Lyon-Albuquerque Phonological Systems Database [@Maddieson2014]. Languages that are not present in either database were classified according to information from the Phonetics Information Base and Lexicon Database [@Moran2019] or other sources from the linguistics literature. A summary of all languages studied here, with further classification details and language-wise sample sizes, is in Table S1.

@@

As in the original tasks cited above, each subtask was presented adaptively via `psychTestR` [@Harrison2020]. To minimize the duration of the experiment, we fixed the length of each subtask at 15 trials, the minimum number of trials with acceptably low mean standard errors, according to the original task designs. Demographic items were presented via `jsPsych` [@deLeeuw2015]. The citizen-science platform distributed the experiments using a modified pre-release version of `pushkin` [@Hartshorne2019]. Readers can try the three music perception tasks at <https://themusiclab.org/quizzes/miq>.

@@

The random-effect structure is particularly helpful for correcting for sampling imbalances, ensuring that no particular languages or countries can dominate the overall effect, while also allowing us to model variation across languages and countries directly (see, e.g., Figure 3).

Note that this analysis approach deviates from our preregistered analysis plan, which involved applying a linear regression model at several sampling levels, without random effects. We made this change on the suggestion of a reviewer and given the utility of mixed-effects models in measuring cultural variation [e.g., @Hilton2022]. For transparency, we report analyses and results from the preregistered approach in SI Text 1.2; these largely replicate the results reported in the next section.

@@

The results robustly replicated the main findings. Speakers of tonal and non-tonal languages could be reliably distinguished on average for both melodic discrimination ($p$ `r permutation_tests$T_NT$mdt$p %>% fp`) and beat alignment ($p$ `r permutation_tests$T_NT$cabat$p %>% fp`) tasks, but not for the mistuning perception task ($p$ `r permutation_tests$T_NT$mpt$p %>% fp`), thus convergently supporting the reliability of our main results.

Last, we used the same linear discriminant analysis approach to compare pitch-accented to non-tonal language speakers (Supplementary Figure 3). Here, the approach failed to replicate the small advantage to pitch-accented language speakers on the beat alignment task ($p$ `r permutation_tests$PA_NT$cabat$p %>% fp`) identified by our mixed-effect modeling analysis (Table 2), but did find a small advantage for vocal mistuning ($p$ `r permutation_tests$PA_NT$mpt$p %>% fp`). We take this mixed result as cause to not interpret either effect for speakers of pitch-accented languages, given both the less clear theoretical basis for such an effect and the noted ambiguity of the pitch-accented language type [@Hyman2009].

@@

First, we tested whether the socioeconomic status of participants was associated with their task scores, as socioeconomic status may vary systematically with both country and native language in our data, and may mediate the opportunities a person has for musical development. For participants who reported living in the United States, we collected additional demographic information. To assess how socioeconomic status may mediate the main findings, we analysed this subset of participants who reported income information ($n$ = `r f(income_n)`). Despite being restricted to the United States, this sub-sample was nonetheless linguistically diverse, including `r f(users$income$Tonal$n)`  speakers of `r f(users$income$Tonal$n_langs)` tonal languages, `r f(users$income$Pitch_accent$n)` speakers of `r f(users$income$Pitch_accent$n_langs)` pitch-accented languages, and `r f(users$income$Non_tonal$n)` speakers of `r f(users$income$Non_tonal$n_langs)` non-tonal languages. We analysed these data using a mixed-effect model of the same structure used in the main analysis, except with an additional fixed-effect term for income. The results show that while income does positively predict performance, such that those with higher incomes tended to perform better, the tonal language melodic discrimination advantage and beat alignment disadvantage held robustly after accounting for these income effects (full statistical reporting is in Table S3).

@@

Our results also help clarify the previously mixed pattern of results concerning the effects of linguistic experience on music processing across different tasks and samples. For example, an advantage for tonal language speakers in melodic pattern processing is consistent with the majority of previous studies [@Alexander2008; @Bradley2016; @Choi2021; @Chen2016; @Swaminathan2021; @Wong2012], though not all of them [@Peretz2011; @Stevens2013; @Zheng2018]. Meanwhile, similar levels of performance on the fine-grained pitch task between tonal and non-tonal language speakers is supported by some prior studies [@Bent2006; @Bidelman2011a; @Stevens2013; @Tong2018; @Zheng2018] but not other studies that point to a tonal advantage/disadvantage on related tasks [@Bidelman2013; @Chang2016; @Giuliano2011; @Hutka2015; @Peretz2011; @Pfordresher2009]. Lastly, while rhythmic abilities in tonal language speakers have been less studied [see @Wong2012; @Zhang2020], a disadvantage in beat discrimination is consistent with recent work showing that tonal speakers give more weight to pitch cues than duration cues; this weighting cuts across auditory domains [@Jasmin2021]. By leveraging a consistent set of tasks and a large sample size, our results make clear that speaking a tonal language has a measurable connection to music perception ability — but not a uniform one, as this linguistic experience produces both positive and negative effects on perceptual abilities of different types.

Why might tonal language experience have these specific effects on music perception? The answer may lie in the shared mechanisms and neural processing resources associated with auditory perception, whether applied to language or music [@Patel2011; @Patel2008; @Kraus2010; @Asaridou2013; @Peretz2015; although see also @Asano2021; @Bidelman2013]. Both tonal languages and music rely on specialized pitch-based sound categories (tone contours or levels in speech; pitch motifs and melodies in music). If these categories are learned and processed through shared, domain-general mechanisms, then improving the efficiency of these mechanisms through practice in either domain should result in mutual improvements [@Asaridou2013 ;@Chang2016; @Delogu2006; @Delogu2010; @Patel2008; @McMullen2004]. But not all is shared. Music relies more on fine-grained pitch structure, even compared to tonal languages, as it is required to support the processing of pitch in the context of a tonal hierarchy [@Albouy2020; @Krumhansl2004; @Peretz2003; @Zatorre2002]. This may explain why tonal language experience did not have an effect on fine-grained pitch processing in music.

Our results do not explain, however, *how* these shared mechanisms might be improved by experience. One possibility is that language experience could shape domain-general perceptual strategies regarding inferences about high-level perceptual categories on the basis of low-level cues: acquired perceptual biases (i.e., from tonal language experience) may aid the processing of some stimuli while worsening the processing of others. In speech, listeners give more perceptual weight to cues that are more informative in discriminating contrasts that are salient in their native language [@Schertz2020; @Jasmin2021] and tonal language speakers rely more heavily on pitch to categorize and produce speech stress when acquiring a non-tonal L2 language compared to native speakers [@Wang2008; @Yu2010].

@@

From the mean and standard deviation, we calculated Hedge's *G* for each task and used it as our effect size measure. We then ran a random-effects model for each category of the tasks. Please refer to <https://github.com/themusiclab/language-experience-music/blob/main/analysis/meta-analysis.Rmd> for commented code of the analysis.

@@

```{r FigureS3}
#| fig.cap = "\\textbf{Figure S3.} To test whether speakers of pitch-accented and non-tonal languages could be distinguished on the basis of their music perception scores, we ran a permuted Discriminant Function Analysis (pDFA). The three histograms show the null distributions for each music perception task, representing the ability to discriminate 'tonal' from 'non-tonal' when these labels have been randomly shuffled, resulting from 10,000 permutations. The dotted vertical line then represents the actual discrimination performance on non-shuffled data. And the p-value is computed as a approximate test from this null distribution and actual score. "
knitr::include_graphics(here("viz", "figures", "FigureS3-1.pdf"))
```
